%! LaTeX
% Revisions:
% 970302: Page layout for various devices... --lct

\documentstyle[12pt]{article}

\voffset= -0.5in %\voffset = -4cm
\hoffset= -0.5in %\hoffset = -2cm
\textwidth = 6.5in  %\textwidth 160mm
\textheight = 9.5in %\textheight 10in
% LaTeX files have a top margin, headheight, and headsep from the 0,0 point.
\headsep = 0pt
%\headheight = 0pt
\topmargin = 0pt

\parindent = 0mm                                % Disable paragraph indenting
\parskip = 10pt plus2pt minus.5pt               % paragraph break

\def\undertext#1{$\underline{\smash{\hbox{#1}}}$}
\def\spell#1#2{\undertext{\bf #1} ({\sl #2\/})\ }


\begin{document}

\title{Chaos User Manual\\V2.3} \author{Sean A. Irvine\and James Conwell}

\maketitle

This is a mythical strategy game for up to eight human and/or computer
players. It is a game of justice and vengeance, of good versus evil,
of life versus death, of light versus shadow, as a group of wizards battle
ferociously to exterminate their opponents and achieve world
domination.

If you dare to compete, you will be faced with beasts of all
descriptions, from everyday animals like horses and lions; to monsters
of hideous proportions, from fire-breathing multi-headed pyrohydras to
extremely dangerous dragons that sail through the air hunting their
favorite food -- you. Creatures from the spirit world are also in
abundance; you may even be confronted with the devil himself. Say your
prayers if you do.

You will experience the most cataclysmic natural disasters ever to
occur in the course of human history. You may be drowned in
swift-moving floodwaters, burned beyond recognition in raging
uncontrolled hellfires, or be swallowed up by the very Earth itself in
horrific earthquakes.

With your own eyes you will witness corpses rising from the dead to
rejoin the battle. You will be struck with terror as your army
succumbs to massive piles of green slime and purple fungi. Even the
humble tree is not always what it seems.

Your army is not restricted to man alone. Your forces will be filled
out with orcs, elves, ogres, and a host of other races. You will
doubtless send many of these creatures to a premature death as they
blindly follow your every command. Only those wizards who use the
strategy of a Perfect General, possess the bloodlust of a Warlord,
make effective use of their Hired Guns and have the zeal to build an
Empire will triumph in the end.

You will utilize and be subjected to some of the most powerful magic,
sorcery, and necromancy known to the human mind. You may be cursed,
drained, and infected with disease. You may disappear into the
ethereal plane or vanish into a timewarp. You will twist time, garble
gravity, and rip holes in space as you criminally violate all the
laws of physics with the unbridled arrogance of the dictatorial
tyrant you really are.

You will cause untold destruction of a once pristine environment as
you destroy entire forests on a whim, saturate the air with
foul-smelling toxic fumes and unleash the apocalyptic fury of nuclear
devastation on all who oppose you.
 
The grapes of wrath never tasted so good.

To survive, you must strike a fine balance between defending your
wizard and attacking the other wizards. Only one wizard can rule the
world, but it can be a long time in coming, with up to eight wizards
chasing the same goal. Furthermore, the wizards may lose control of
their powerful magic and suffer the dreaded fate of mutual
annihilation; forever cleansing the world of disruptive magic!

\section {System Requirements}

Versions 2.3d and below may be played on any Amiga with 1 megabyte of ram
and OS 1.2 or higher.  Yes, even if you have an old 1985 A1000 with Kickstart
1.2 you can play this great game so long as you have 1 meg or more of ram.

The authors went to great lengths and much trouble to ensure compatiblity
with all the old Amigas.  We hope you enjoy the fruits of our labors.

Version 2.3d is the LAST version that will work on any Amiga with 1 meg of
ram.  ALL versions above 2.3d require at least 2 megs of ram, a hard drive,
at least OS 1.2 or higher and a Motorola 68000 CPU or higher.

2.3d = demo version.  2.3f = full version which only registered users can
obtain.  The demo version cannot load games, and stops at turn 18.


\section {Installation}

Just toss it on your hard drive somewhere.

This game requires that req.library be in your LIBS:
Its probably already in there, but if it isn't then we included it for you
in the libs directory of this archive.  Just copy it into LIBS: and you
are all set.

There are no other files that you have to mess with.


\section {The Real Stuff}

It should be unnecessary to read all the instructions to play the
game.  For those who are itching to start skip to Section~1.1 to see
how to load the game and then just follow your nose. For the rest of
you, the following describes how the game is played.

Each player (human or computer controlled) starts the game with a
wizard and a repertoire of spells. The idea is to be the last wizard
alive and to amass the greatest score. Once your wizard is killed you
are out of the game, so protecting your wizard is vital to your
survival. The options available at the start of the game allow for a
wide range of scenarios.

A game is organized as a series of rounds. Each round consists of
several stages, and each living wizard may partake in each stage of
each round.  The major stages are: spell selection, spell casting, and
movement.  Each stage is described in detail below. As well as these
main stages there are also a number of auxiliary stages performed
solely by the computer.

The game can be played almost entirely with the mouse and is icon
driven.  All the entities in the game have animated graphical
representations.  A text region at the bottom of the screen is used to
display other information.  Initially, you may be perplexed by the
wide range of spells available but after a few games you should become
comfortable.

The world once was comprised of a 7000 mile diameter sphere covered
with beautiful vegetation, crystal clear water and wonderful wildlife.
But that was before the 1000 year wizard war of Irvine vs.~Conwell;
the culmination of which resulted in the destruction of most of the
prime material plane of existence.\footnote {Human Beings commonly
  refer to ``The Prime Material Plane of Existence'' as ``The
  Universe''.  Other important planes of existence are the Astral and
  Ethereal planes of existence.} All that is left is a flat
rectangular board of square cells, measuring 17 across by 14 down.

The game is played on this rectangular board of square cells. You move
your pieces around the board using the mouse. The cost of a move
depends on the distance between the initial cell and the final cell.
It costs more to move diagonally than it does to go vertically or
horizontally.

It is important to understand there is a distinction between the
gaming environment supplied by the computer and the computer
controlled entities.  When a computer player moves it does not get
access to any information that you cannot get. In fact, the computer
bases it moves on less information than a human player.


\subsection {How to Win}
\begin{itemize}
\item Kill all enemy wizards.
\item Kill or block all independent generators.
\end{itemize|

A generator needs an empty space on which to generate creatures onto.
So if you can block all the spaces surrounding a generator, then it cannot
generate anything.


\subsection {Game Startup}

The game can be started from the Workbench or from the command line.
Some of the options can only be controlled from the command line. The
usual way of starting from the command line is to enter the command
{\tt chaos -t} when in the directory containing the chaos files. There
is an icon for starting from the Workbench. Installation is discussed
in Section~\ref{install}. You should ensure Chaos has a stack of at
least 50000 bytes. From the command line this can be done by typing
{\tt stack 50000} before starting chaos. Those who have received Chaos
on a self-booting disk need not worry about this matter.

These instructions are accurate for 2.3 of the game. However, Chaos
is a rapidly evolving program and new features are continually being
added and some instructions may become obsolete or inaccurate in the
future. Only registered users are guaranteed of being informed about
new releases.

After passing the introductory screens you will arrive at the first
setup screen. This screen allows you to select which wizards will be
playing, whether or not they will be computer controlled, the teams,
and their names. By default it starts with two human controlled
wizards on different teams. The `T' column is used for team selection.
The icons appearing in this column represent a team. By clicking on
them they move on to the next team in the sequence. If two wizards are
showing the same icon this means they will start on the same team.
There also exist spells which can alter the teams during the course of
a game.

Once you are satisfied with the setup click on `Start' which will take
you to the second initialization screen. On this screen you can
select the number of generators (default 4, maximum 99), number of
spells (default 20, maximum 99), and level of computer wizards
(maximum 9).

The wizard level influences only the allocation of spells to computer
players and does not influence the way the computer moves or casts
spells. Any wizard can improve their level rating by casting Level
spells during the game. Doing so increases the chances of getting
useful, rare and powerful bonus spells.

You should decide how many turns the game should last for. The default
action is forever which means the game continues until all human
players have been killed or one human player remains and all computer
players are dead. Alternatively, any number up to 999 can be
specified.  The number of turns remaining is displayed at the top
right of the screen as a series of tally marks. If these are shown in
yellow then you are into the last few turns.

You can also turn ``Texas Trash'em'' on or off.  Texas Trash'em allows you
to optionally trash (discard) one of your spells in addition to casting a
spell each turn.  This allows you to discard spells you don't like or
don't want.  In version 1.8 and below, computer players do not discard
spells.  In version 1.9 and up you MUST discard a spell each turn but
casting a spell is optional (unless there are 5 spells or less left in your
spell list in which case you don't have to discard if you don't want to.)

By default the game will operate in ``Classic'' mode. In this mode
only a limited number of spells are available. This mode is
particularly suited to neophytes. If you switch off this mode then the
full repertoire of spells is made available.

You may also select the number of artifacts that will be randomly placed
on the board. Allowable settings are 0 to 99.

Artifacts are like a regular creature object except it belongs to
NOBODY until someone moves adjacent to it.

An Artifact's owner is determined at the beginning of each movement phase.
It is determined according to who has the most creatures adjacent to it.
In the event that 2 or more players have an equal number of creatures
adjacent to it then it remains NOBODY's.

Inanimate objects and other artifacts do {\em NOT} assist in controlling
artifacts.

Currently there are only 4 artifacts:
Bolter Wall Artifacts
Dalek Artifacts
Robot Artifacts
Stone Golem Artifacts

Artifacts have 0 movement points.

Most magic spells can be cast on Artifacts so you can heal an Artifact
or increase its capabilities, etc.

If you play Chaos against human opponents who hide in a corner, cowering in
fear, then try playing the game with some artifacts.  The sight of a
nice, big, powerful artifact just sitting there waiting for someone to
claim it can be a strong incentive for them to charge out after it and
has been known to spark intense power struggles.

Hidden Artifacts:
If this is turned on then artifacts will be buried underneath
walls, trees, rocks and other inanimate objects.  These objects will have
to be destoryed in order to take posession of the artifact underneath.


You can also select the number of scrolls that will appear on the board.
Allowable settings are 0 to 99. A scroll is a spell lying on the board
which is free for the taking. Isn't that cool? Scrolls are magically
inscribed on extremely delicate, dry, brittle, paper. They are
therefore excessively fragile and even the tiniest amount of damage will
destroy them. The scroll rules should be set to one of the following:

\begin{enumerate}
\item Scrolls can be picked up by anybody.
\item Scrolls can be picked up by wizards only.
\item Scrolls can be picked up by anybody except flying creatures.
\end{enumerate}

Inanimate objects can never pick up scrolls.

There are normally only 3 ways to affect scrolls:
\begin{itemize}
\item By entering a cell that contains a scroll and surviving there until
   the end of the complete round. The complete round means until
   everyone has finished their turn for this round, including all
   wizards and independents.  If you survive in a cell containing a
   scroll until the end of the round then the scroll is magically
   added to your commanding Wizard's spell list.
\item Destroy them by shooting them.
\item Destroy them by Burying them with the Bury spell.
\end{itemize}

 
NOTE: Shooting the scrolls of Undead Elementals may produce Ungood side
effects. }8>
 
If you play Chaos against human players who just sit in a corner and
avoid conflicts with each other then try playing the game with some
scrolls laying around.  This can draw them out of their corner and
start up some fierce battles!


Scroll Delay:
Allowable settings are 0 to 6.

This controls how many consecutive turns you must control the cell before
you are awarded the scroll.  0 means you will acquire it at the end of
{\em this} round.  1 means you will acquire it at the end of the
{\em next} round and so forth.

If at the end of a round, you no longer control the cell then you must
start over again so if Scroll Delay is 6 and you lose control of the cell
after only 5 turns then you must regain control for 6 consecutive turns to
obtain the scroll.

You are permitted to move creatures into and out of the cell containing the
scroll without any penalty so long as you control the cell with a
scroll-capable creature at the end of the round.


Hidden Scrolls:
If this is turned on then scrolls will be buried underneath
walls, trees, rocks and other inanimate objects.  These objects will have
to be destoryed in order to take posession of the scroll underneath.


At this time you should not change the board size.

When you have made any desired changes, click on `Start' to begin the
game.

\subsection {Spell Selection}

For each human wizard in turn, the right of the screen is filled with
the currently available spells for that wizard. Information can be
obtained about the spells by reading about the spells in this manual
or by selecting the `?' gadget and then clicking on the spell you wish
to know about.  Or simply by pointing the mouse at the spell in question
and then pressing the `q' or `?' key on the keyboard.

A wizard can decide not to choose a spell by clicking
the return gadget, whence control will pass to the next wizard.

Assuming you wish to pick a spell, simply click on the desired spell.
During the casting stage you will attempt to use your selected spell.
Care should be taken in spell selection. Consider such things as the
range of the spell, the situation of the wizard, your position on the
board, and whether line of sight is required for the spell.  Just press `c'
(for Casting possibilities) while the mouse is over a spell and all the
possible places that you could cast that spell will be highlighted on the
board.

Once selected, a spell must be attempted.  If the spell is aborted, it will
still be considered used and will not appear on your spell list in the next
round.

If ``Texas trash'em'' was selected then you will also be required to
select a spell to be discarded (unless you have less than six spells
remaining). Such a selection must be made even if you choose not to
select a spell for casting.

\subsection {Spell Casting}

During spell casting, each player who selected a spell during the
spell selection stage will in turn get a chance to actuate the spell.
Computer controlled wizards will attempt to cast a spell at their
position in the sequence. In addition, each generator on the screen
might produce a creature. Certain other creatures also occasionally
cast spells.

For each player in turn the spell selected is shown on the bottom line
of the screen, together with the number of attempts and the range for
the spell. For example,

\centerline {\tt Gray Elf x1  Range=2}

means 1 Gray Elf is being cast, and the elf can be cast either
immediately adjacent to the wizard or one cell removed from the
wizard (just press {\tt c} to see all possible legal cells).
A range of zero means the spell can be cast simply by pressing
the left mouse button and there is no need to select a cell. Such
spells are often special powers for the wizard. The computer
automatically applies the spell to the correct wizard or creature.
Certain spells have a range of 15, these spells can be cast on any
appropriate cell on the screen and do not require line of sight.

Most spells have a fixed range, measured in a straight line from the
casting wizard. For instance, nearly all creatures have a range of one
and so can only be cast immediately adjacent to the wizard. All these
spells require line of sight. The line of sight is measured from the
centre of the wizard's cell to the centre of the destination cell.
You can see over blank space, dead creatures, pits, pools, and fires.
Sometimes line of sight can be obtained by waiting until an
appropriate time in the animation sequence.\footnote{To experienced
  players this is known as {\bf forcing.}} By pressing {\tt c} you can
ask the computer to show you all legal target cells.

Spells can be aborted by pressing the right mouse button. Aborted
spells still get used up, so try to avoid having to abort spells. If
you find yourself aborting spells often, you are probably not choosing
spells appropriate to your situation.

Most spells can be cast only on certain types of cell (the program
will notify you of any errors you make in selecting the cell). If you
make an error you are free to choose a new cell. For instance,
creatures can be cast on empty cells or cells containing dead
creatures, but cannot be cast on other creatures or inanimate objects.

After each wizard has cast, the generators may produce creatures on
any cell adjacent to the generator which is empty, a growth, or
contains a dead creature. Adjacent cells occupied by creatures or
inanimate objects will never be generated upon.

\subsection {Board Update}

The board update is handled entirely by the computer. The update
carries out activities which are not directly under any player's
control. During the update any or all of the following things might
happen:
\begin{itemize}
\item The growths (earthquakes, fire, flood, fungi, orange jelly,
  gooey blob, and green ooze) are permitted to grow. Each cell
  containing a growth may or may not grow. If a cell is going to grow
  it will do so in only one direction in any particular turn. The
  direction of growth is chosen at random. Earthquakes cannot `grow'
  diagonally. Some of the growths grow faster than others, orange
  jelly is particularly voracious while floods tend to be sluggish. If
  a cell doesn't grow it might dissipate instead. You may have some of
  your creatures buried, killed, or revealed in this growth process.
  Should your wizard be buried or killed you are considered dead and
  your game ends (an exception occurs in the case of a wizard having
  previously cast Lich Lord).
  
\item Any tempests on the board may move. A creature in the path of
  the tempest will be thrown to a new random location. On very rare
  occasions a creature will be killed as a result of the throw.
  
\item Any vortices may move. A creature in its path will disappear
  into the future.
  
\item Ropers get tired of constantly sitting in the same place.
  Sometimes they decide to teleport to a new location.
  
\item There is a chance that each magic castle and dark citadel on the
  board might collapse. If a wizard is inside the castle or citadel
  when the collapse occurs the wizard will be granted a bonus spell.
  
\item Any magic wood occupied by a wizard might collapse, resulting in
  a new spell for the wizard.
  
\item Each sleeping entity is examined and has a chance of dying or
  collapsing. The weaker things are more likely to be destroyed than
  the stronger things. The destruction will be complete and no trace
  of the entity will remain.
  
\item Each Haunt has a small chance of changing sides.
  
\item Each Spriggan has a small chance of falling asleep.
  
\item (When impurities are turned on) With a small probability an
  object might decide to change sides. This may work to your advantage
  or disadvantage.  It is even possible that a generator will become
  your friend and decide to emit creatures for you.
\end{itemize}

\subsection {Movement}

The movement stage is the most complex stage, since normally there are
a large number of creatures to be moved and each creature has a number
of possible moves. Each player is given a chance to move in turn. The
strategies employed during the movement stage will have the greatest
effect on your survival in the game. Movement involves not only the
shifting of creature positions but also combat.

Dead creatures, sleeping creatures, and growths cannot be moved. No
inanimates can be moved, although shadow wood trees and bolters may be
used to attack enemy creatures\footnote{Unlike creatures, most trees
  cannot attack other inanimate objects, or growths} in adjacent
cells.  Special objects like ropers and wasp nests are controlled by
the computer.

A player is able to move each creation once in each turn. To end your
turn click on the return gadget and control will pass to the next
player.  Players having cast the Move It spell in the current round
will get a second chance to move after all other players have moved,
but before the independents move.

A creature is said to be {\bf engaged\/} if it is adjacent to an enemy
creature and is unable to retreat. Engaged creatures cannot be moved
but may still be able to fight. How likely a creature is to be engaged
is dependent on its manoeuvrability and the number of adjacent hostile
creatures.

To move a creature simply click on the creature using the left mouse
button. Normally, the distance the creature may be moved will be
displayed in the text bar. Depending on the creature, it will be
flying movement or walking movement. There is a special cursor when
flying movement is active. To see all legal moves press the {\tt m}
key.

\subsubsection {Flying Movement}

For flying creatures you can pick any cell within the given range
except cells occupied by your team's creations. If the cell is empty
or contains a dead creature then the creature will move to the
selected cell, otherwise an aerial attack will be made on the selected
cell.

\subsubsection {Aerial Attacks}

An aerial attack results when a flying creature is moved onto a cell
occupied by an opponent. The flier will attack the current occupant.
If unsuccessful the attacker will remain at its current location. If
successful the current occupant will die and the flying creature will
be moved into the cell.

\subsubsection {Walking Movement}

A ground based creature can only move one cell at a time and you must
show the computer exactly how to move multiple cells by clicking in
each cell in turn. Many walking creatures can only move one cell
anyway.  For those which can move more than one cell you should select
a sequence of adjacent cells---you can abort the move at any time by
pressing the right mouse button. During the course of a move a
creature may become engaged in which case it will move no further.

\subsubsection {Combat}

Engaged creatures can take part in combat. You cannot attack your
team's creatures (but see ranged combat). Living creatures cannot
(normally) attack undead creatures, otherwise anything goes. Once
engaged, an attack is induced by clicking on the cell containing the
creature you wish to attack.  If successful, the defendant will be
killed and the attacker will move onto the corpse. For a successful
kill of a strong opponent you may be awarded a bonus spell.

It is possible to attack inanimate objects and growths, although you
cannot get engaged to them. Be wary of killing growths with buried
creatures beneath, as the buried creature will be revealed and will
probably be quite hungry.

\subsubsection {Ranged Combat}

Some creatures such as elves, dragons, and wizards carrying bows, have
additional attacks. Such an attack is indicated by the {\bf sights
  pointer}. Line of sight is required for most of these weapons. It is
possible to shoot your own creations or creations of your team
members, so choose your targets carefully. Occasionally it might be
desirable to kill one your own creatures due to insanity or disease
(especially if you are playing for points).

Creatures which posess the {\bf Archery} skill (such as the Wood Elf and
Gray Elf) do not need line of sight to shoot a target.  Their shots simply
arch over any intervening obstacles. 

\subsubsection {Wizard Movement}

Under normal circumstances the wizard can be moved like any other
creature.  However, wizards are capable of special movements. Wizards
can climb magic trees, enter castles, and ride certain creatures. For
convenience these activities are collectively called {\bf mounting}.
You cannot mount if you are engaged, otherwise mounting is effected by
simply moving the wizard onto the cell occupied by the entity you wish
to mount.

Dismounting is somewhat more complicated as the computer needs to
distinguish which of the two entities is to be moved. For trees and
castles the computer will assume the wizard. For ridden creatures the
default is the creature. To dismount from a creature the following
sequence should occur:

\begin{itemize}
\item Select the creature the wizard is riding.
\item Click on the `D' gadget at the top right of the screen (which
  will now be unghosted).
\item Move the wizard to a cell adjacent to the creature (wizards with
  high movement allowance or wings still only move one cell when
  dismounting).
\end{itemize}

Mounting is generally a good idea, as it hides the wizard from view
and gives the wizard additional protection.

\subsubsection {Ending Your Turn}

Once you have moved all creatures that you wish to, select the end of
turn gadget. Computer controlled wizards will move their creations at
the appropriate point in the sequence. After all wizards have moved
the independent creatures will be moved by the computer according to
the same rules. Finally, any wasp nests, ropers, and fungi on the
screen will attack adjacent opponents.

The round is now effectively over and the game returns to the spell
selection stage for the next round.

\subsection {Scoring}

The computer keeps track of a score for each player and this is
indicated by the computer displaying a ranking of all wizards and the
independents at the end of the game. There is a score and a life
column. The life column is the sum of the life force of a wizard's
creations excepting walls, castles, and certain other inanimate
objects. The current scores can be examined during the spell selection
stage using the menus.

Points are awarded for killing enemy creatures and wizards. The
stronger the creatures you kill the more points you get. Bonus spells
are awarded for killing particularly strong creatures. Points are also
awarded for successfully casting some difficult or unusual spells.
Points are even awarded for destroying or burying corpses.

\subsection {Information}

The `?' gadget can be used to get information about objects on the
board (at any time) and about spells (during spell selection). For an
object on the screen, a window is displayed showing various
characteristics of the chosen object. The `?' gadget will not reveal
any useful information when applied to a cloaked creature, unless you
own it.

The information screen for a wizard or a cell containing a mounted
wizard will also display the various power-ups held by the wizard as
icons.

Most creature's Combat stats affect their opponent's Life stat but
Some creatures, like the Mind Flayer, attack Intelligence or some
other stat. Sometimes a creature's Combat or Ranged Combat or Special
Combat will affect more than one of your Enemy's stats at a time.

Just press the Left Mouse button on either Combat, Ranged Combat or
Special Combat and all the stats that it attacks will begin flashing.

\subsection {Keys}

By pressing the number keys {\tt 1}--{\tt 9} the creations of the
corresponding player can be highlighted. Key {\tt 9} highlights the
independent creatures.  Creatures who have not moved are highlighted
in a cycling color.  Creatures who have moved are highlighted in red.
Allies are highlighted in blue.

Pressing {\tt q} for Query is another way to bring up the
information screen for a creature.  Once brought up you can press {\tt q}
again to quit the info screen or press any other key to continue on to the
2nd info screen.

Pressing {\tt c} for Cast during the spellcasting phase will show all the
legal targets for the spell about to be cast.  It takes into account all
the rules for spell casting such as range, line-of-sight, wizard powerups
and the exact type of thing in the target cell (growth, creature,
inanimate, etc.)

Pressing {\tt m} for Move shows all the legal moves that the currently
highlighted creature can make or if no creature is currently
highlighted then it shows the legal moves for the creature under the
mouse pointer.  `Moving' includes `attacking' so often the cells of
enemy creatures will be highlighted (because it is legal for you to
attempt to move there) even though you perhaps don't have a snowball's
chance in hell of really killing the creature (e.g. orc vs. stone
giant).

Pressing {\tt r} for Range shows all the cells that are within range of
the currently highlighted creature's ranged weapon or if no creature is
currently highlighted then it shows all cells that are in range of the
creature under the mouse pointer.  {\em Note\/}: this shows all the
cells that are within range not all the cells that you can actually
hit.  Many times the cells shown will be unshootable due to
intervening objects blocking line-of-sight or the fact that undeads
cannot be shot by living creatures.

Pressing {\tt s} for Shoot shows all the cells that can actually be shot
by the currently highlighted creature's ranged weapon or if no creature is
currently highlihgted then it shows all the cells that can actually be
shot by the creature under the mouse pointer. {\em Note\/}: this takes
into account line-of-sight and living/undead restrictions.

Pressing {\tt l} for Line-of-Sight shows all the cells that can be seen by
the currently highlighted creature or if no creature is highlighted then it
shows all the cells that can be seen by the the cell under the mouse pointer.
This is particularly useful when you are thinking of casting a spell that
requires line-of-sight and you want to see if your wizard can actually see
the intended target.  It is also good for checking out where enemy wizards
can/cannot see to so you can hide your creatures behind walls and things to
protect them from subversion/betrayal/abduction/sleep/braindrain etc.

Pressing {\tt $\langle$shift$\rangle$} in combination with {\tt m},
{\tt r}, {\tt s} or {\tt l} will {\it always\/} show the relevent cells
for the creature under the mouse pointer, regardless of any currently
highlighted creature.

\subsection {Menus}

The menus can only be used during the spell selection stage. They are
primarily concerned with loading and saving games, but it is also
possible to turn the speech on and off. The sound option is used to
switch off the sound while the independents move. Turn it off if you
want the game to go faster but be warned it is hard to follow what is
happening.  There is also a menu option to view the current scores.
The ShowCast menu item is used to control the graphical effect used
during spellcasting.  Turin it off if you want the game to go faster
but be warned it is hard to follow what is happening.

\section {\label{install}Regulations and Installations}

 The source code, documentation, levels and executables
are copyright \copyright Sean A. Irvine, James Conwell 1994,
1995, 1996, 1997, 1998, 1999 and may not be modified or distributed
without the author's permission.

Chaos will work under at least Workbench 1.2, 1.3, 2.04, 3.0 and 3.1 and
should work on anything subsequent. It requires an Amiga supporting PAL mode.
All Amigas manufactured since 1989 support PAL mode.  Any Amiga
manufactured before 1989 supports PAL mode if it has been upgraded with
an enhanced Agnes chip. (Enhanced Denise not needed).

If you live in the USA or Canada then you'll need to boot your computer
into PAL mode.  The best way to do this is to use the program "degrader",
though in this case you aren't going to actually degrade anything.
With degrader select "50 Hz", "50 Hz System" and "Survive Resets" and
your Amiga will be rebooted in PAL mode and will think it is in a European
country operating in PAL mode.

Normally a North American Amiga operates in NTSC mode 640x400 resolution.
PAL mode produces a resolution of 640x512.

Chaos makes use of the {\tt narrator.device} and {\tt req.library}.
As of version 2.2 Chaos no longer makes any use of {\tt ReqTools.library).

If the {\tt-t} option is used the {\tt translator.library} is also used.
Other zrequired files are the executable {\tt chaos} and the data files
{\tt pool.ssc} and {\tt chaos.prb} all of which must reside in the current
directory when Chaos is started.  There are various other fonts, music and
SoundEffect files which reside in the directories of the same name within
the current directory.

The {\tt -t} option forces translation of player names for the speech
and requires the {\tt translator.library}. A few other options are
supported, to find out about these use {\tt chaos -h} and then press
$\langle\cal A$-M$\rangle$ to return to the CLI screen.

\section {Future Additions}

We can only work on Chaos in our spare time but are always pleased to
hear suggestions for improvements from people who play the game.

Here is a list of features that may be added in the future:

\begin{itemize}
\item More spells (send me your favourite ideas, and pictures of
  wanted creatures).
  
\item Improved sound and sampled speech.
  
\item More flexibility in game options, such as allowing players to
  buy spells, and so on.
  
\item Level editor.
  
\item Amiga guide documentation.
  
\item HTML documentation.
\end{itemize}

\section {Known Bugs}

Chaos has been extensively tested with {\sl Enforcer\/} and in our
experience does nothing nasty. However, crashes may happen, if they do
it is most helpful to me if you can note carefully what was happening
when the crash occurred. There are a few less serious bugs within the
game which may be fixed in the future.

\begin{itemize}
\item {\sl Bolter Bug\/}: When a computer controlled Bolter gains a
  movement rating (e.g. by the Speed spell) it is able to move. A
  human controlled Bolter does not have this ability. Since this is
  rather amusing when it happens, we have no intention of removing it.
\item {\sl Reinstatement Bug\/}: When Dark Power, Vengeance, Decree,
  or Justice succeeds in killing a creature upon a corpse, the
  creature is killed but the corpse is brought back to life. This ``bug''
  is deliberately incorporated in the program.  Advanced players will
  exploit this fact in the obvious way.
\item {\sl Cat Lord Bugs\/}: Theoretically there should only ever be
  one Cat Lord. However, it possible to create more than one by a
  variety of methods: (i) using Alter Reality, (ii) using mutate,
  (iii) by waking a sleeping one, (iv) by raising a dead one, (v)
  using Request. Multiple Cat Lords will not cause the game to crash
  but may result in cats changing sides several times, moving multiple
  times, and other undocumented effects. We leave it to you to explore
  the possibilities.
\item {\sl Classic Bugs\/}: The classic mode is not perfect and it is
  possible that spells not normally available will occur. Known cases
  are the faun casting sleep.  The thundermare and iridium powerups
  are still possible. If generators are used then the bipedal generator
  will still generate creatures not normally available in classic mode.
\end{itemize}

\section {Registration}

If you register this game then you will automatically be an Official
Playtester and your name will be placed in the credits.  This means you'll
be famous once the Playstation, N64, DreamCast, Mac and PC versions are
released.

If you would like to register this fine game and recieve the full version
with loading, saving and unlimited turns capability then please send $35.00
US to:

If you are in North America, South America, Europe or Scandinavia and are
sending US $ Cash (That's actual money, not a check or money order) then
send it to:

\begin{tabular}{l}
James Conwell\\
16806 Carrack Turn\\
Friendswood, TX 77546-2311\\
USA
\end{tabular}

{\tt JamesConwell@yahoo.com}

Due to the lameness of the US banking system, European checks and currency
are nearly worthless here so _please_ send actual US dollars.


If you are in New Zealand or Australia, or if you insist on sending foreign
currency then please send it to:

\begin{tabular}{l}
Sean A. Irvine\\
5 Westra View\\
Tawa\\
Wellington\\
New Zealand
\end{tabular}

{\tt sairvin@xtra.co.nz}

Apparently the NZ banking system isn't as arrogant as the US system so they
apparently don't have the same problems with foreign currency as we have in
the USA.


Be sure to tell us if you want your registered version sent to you via
Email or snail mail.

\section {Equipment Used}

The following hardware and software was used during the development of
Chaos:

\subsection {Hardware}
Amiga 3000 with 25Mhz 68030, 225 Megs HD, 14 Megs RAM and 1950
Multisync Monitor for coding.
Amiga 2000 with 9 megs ram for testing 68000 compatability.
Amiga 1200 with 2 megs ram and 120 meg HD for testing AGA compatability.
Hewlett-Packard Laserjet 4M LaserPrinter with 6 megs ram for printing out
long program listings and documentation.
Perfect Sound Digitizer, Radio Shack Mixer, Alesis 3630
Compressor/Limiter, Telspec Multiline Teleconferencing Telephone
System.

Beginning July 13, 1998 a 4.3 gig hard drive began to be used.

\subsection{Software}
SASC 6.5, Cygnus Ed Professional 3.0, Devpac Assembler 3.14,
A68K Assembler, Audition 4, Protracker, JR-Comm,
Octamed Sound Studio Professional, Flashfind, Add36k, True Basic 2.0,
YAM 1.3.5, Miami 3.0, Ibrowse 1.2, AmFTP 1.5x, AmIRC 1.6
and lots of other stuff.
 
\subsection{Miscellaneous}
James Conwell has had the following utilities installed on his Amigas for
years and has used them every day and they work great and he recommends
them to everyone: CPUblit 1.0, Arq 1.78 (Animated Requesters), FastExec
(Loads Exec.library to fastram instead of chipram; speeds up multitasking
and interrupts, etc.)

\section {Acknowledgments}

Many people have made useful and not so useful suggestions about
Chaos.  Whenever possible we try to incorporate the suggestions of
ardent fans.  While most of the correspondence we receive is strongly
supportive, we are open to criticism as well. Naturally, we give
considerably more weighting to the suggestions and requests of
registered users, but will gladly listen to the opinions of anyone.

The following people have helped in many ways and Chaos would
definitely not be the program it is without their support (our
apologies for any glaring omissions): Mike Baron, Julian Gollop, Ingo
Holewczuk, Callum Irvine, Glen Irvine, James Levell, Nicole More,
Richard Romanowski, Stephen Smith, Michael Van Elst, Accolyte, Mikael
Kalms, Kalsu, Bryan Ford, Matthew Smith, Jonathan Rudolph, Joseph Conwell,
Jerry Conwell.

Req.library by Colin Fox and Bruce Dawson.

\TeXnical Assistance by Lance Tagliapietra.

32-bit Multitasking Operating System by Carl Sassenrath.

Hardware Design by Jay Miner {\em et.  al.}

\section{Technical Features}
This game was coded in a mixture of JC + Assembly code.  We use JC for the
main game and super high speed optimized assembly code for certain
time-critical things like the music and the ray tracing and certain math
operations and certain graphics operations.

Version 2.3 contains 313 spells.

If you have an 020 or higher processor then
this game features 50 fps animation in the quadrascopes and
quadspectrograms.  If you only have a 7 Mhz 68000 then you will only get
about 25 fps.

The sprite that highlights the current creature only runs at 25 fps no
matter how fast your computer is.  We did this on purpose because when it
moves faster than 25 fps it becomes unpleasing to the eye.

If you have an 020 or higher processor then the color cycling on the QUIT
and NEW GAME requesters moves at 50 fps.  If you only have a 7 Mhz 68000
then it is only about 25 fps.

The vector graphic pentagrams on the ABOUT requester rotate at 50 fps
on all Amigas.

The intro music utilizes 4 simultaneous channels of audio, each one at up
to 28,000 samples per second.

The introductory credits are displayed in 640x512 resolution with 8 colors
and smooth full-screen scrolling.

This game utilizes ray-tracing.  When calculating line of sight to a target
a beam of light is traced from the source to the target.  If the beam gets
through then you have line of sight.  If the beam is blocked then you do not
have line of sight.

Ray Tracing is an inherently slow process so if you only have a 7 Mhz 68000
you will notice a delay of up to .5 seconds when you push the ``l'' key to
see all the cells on the board that you have line of sight to while the
computer ray-traces all the possiblities.  If this is too slow for you then
please upgrade to a 25 Mhz 030 and then you'll be happy 8)

This game was programmed to be 100% OS friendly and Multitasking friendly,
however it will not work on graphics cards because
1) Most graphics cards are lame and don't have sprites.
2) This game contains certain high speed routines that write directly to
   your display memory in order to go 10x faster than using slow OS calls.
   This is totally legal and system friendly; it just isn't compatible with
   add-on graphics cards, so if you have a graphics card you'll need to
   play the game on your regular built-in chipset instead of the graphics
   card.
3) Neither of the authors owns a graphics card.

This game can be played on Amigas which are equipped with graphics cards
however no benefit will be derived from the presence of a graphics card.
It didn't make sense for us to make graphics that go 10x slower for regular
Amiga owners just so that those with graphics cards could go a bit faster.
Neither of the programmers of this game own a graphics card.  If you want
us to make a new ``graphics-card-aware'' version of the game then you'll
have to buy us a graphics card to work with.

Version 2.3 utilizes a main gameboard resolution of 320x256 pixels with 32
colors.  The text area at the bottom of the main screen is displayed in
640x256 resolution with 8 colors.  The information panels are also
displayed with 640x256 resolution in 8 colors.

Bill Gates does not profit in any way from the sale or use of this game.

\section*{Appendix: Spell Descriptions}

To find out about spells when playing the game the `?' gadget can be
used. Not all spells are equally likely to occur and some can only be
obtained as bonus spells or by meditation.

Chaos~1.5 and up features a file called {\tt chaos.prb} which contains the
probabilities for the various spells. These can be changed using a
text editor. Please read carefully the instructions in the file if you
wish to do this, and remember to save a copy of the original file.

\input {spells.tex}
\end {document}
